Verification of concurrent design patterns with data

### 2019-04-25 2018-Arch-pNets/Coordination Camera ready/arch-pNets.tex
Coordination revised: some improvements

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\Eg}[1][\ ]{E.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}
\newcommand{\resp}[1][\ ]{resp.#1}

@@

%% \newcommand{\order}{\leqslant}
\newcommand{\order}{<}
\newcommand{\ordbool}{\ensuremath{\sB^{\order}}}
\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\signature}{\ensuremath{\Sigma}}
\newcommand{\variables}{\ensuremath{\cV}}
\newcommand{\Talg}{\ensuremath{\cT_{\signature,\variables}}}
\newcommand{\actions}[1]{\ensuremath{\cA_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\monexprs}[1]{\ensuremath{\sE^{\order}_{#1}}}
\newcommand{\boolexprs}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\guards}[1]{\ensuremath{\ordbool_{#1}}}
\newcommand{\assigns}[1]{\ensuremath{\sA_{#1}}}
\newcommand{\updates}[1]{\ensuremath{\sA^{\order}_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\data^{#1}}}
\newcommand{\val}[3][]{\ensuremath{#1{\sigma}^{#2}_{#3}}}

%% CTL
\newcommand{\AG}[1][\ ]{\ensuremath{\mathtt{AG}#1}}
\newcommand{\AF}[1][\ ]{\ensuremath{\mathtt{AF}#1}}
\newcommand{\EF}[1][\ ]{\ensuremath{\mathtt{EF}#1}}
\newcommand{\EX}[1][\ ]{\ensuremath{\mathtt{EX}#1}}
\newcommand{\AW}[3][\ ]{\ensuremath{\mathtt{A}\left[#2\ \mathtt{W}\ #3\right]#1}}
\newcommand{\EU}[3][\ ]{\ensuremath{\mathtt{E}\left[#2\ \mathtt{U}\ #3\right]#1}}

@@

\begin{abstract}
We provide a solution for the design of safe concurrent systems by
compositional application of verified design patterns\mdash 
called \emph{architectures}\mdash to a small set of functional
components.  To this end, we extend the theory of architectures
developed previously for the BIP framework with the elements necessary
for handling data: definition and operations on data domains, syntax
and semantics of composition operators involving data transfer.  Under
monotonicity assumptions on guards and assignments, composition of
architectures preserves their characteristic safety properties.
To verify that individual architectures do enforce their associated
properties, we provide an encoding into open pNets, an intermediate
model  that supports SMT-based verification.
%% Additionally, the translation into  open pNet  \noteEM{ICI}{gives
%% us a strong insight and a comparison of the coordination mechanisms
%% provided by the two frameworks}.
The approach is illustrated by a case
study based on a previously developed BIP model of a nanosatellite
on-board software.

@@

We assume the existence of a universal data domain
given as a partially-ordered set $(\data, \order)$, potentially
encompassing several copies of any given data type with
different orders.  We assume that $(\data,
\order)$ comprises both the unordered set of Booleans $\sB =
\tuple{\{\true, \false\}, \emptyset}$ and the naturally ordered one
$\sB^< = \tuple{\{\true, \false\}, \{\false < \true\}}$, and
similarly for integer and real numbers; as well as the set of
intervals ordered by inclusion.
%In practice, and in the examples of
%this paper, all the variables are sorted, meaning that their values
%can only range over given subsets of $\data$, such as $\sB$ or
%$\sB^<$.
%
When speaking of an ordered sort, \eg $\sB^<$, we will assume that it
forms a meet-semilattice and denote by $\wedge$ the meet
operator.

@@

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
%\begin{itemize}
%\item
$S$ is a set of states; %$\vars(s)$ denotes the set of associated variables;
%\item
$s_0 \in S$ is the initial state;
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
%\item 
$\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
% SB: I removed a new paragraph here, ok?
$\langle \alpha,~g,~u\rangle$,
where $\alpha$ is a parameterised action, $\alpha \in\actions{\variables}$; 
$g\in\boolexprs{\variables}$ is a guard over variables of the source state and the 
action, and $u\in\assigns{\variables}$
assigns updated value for variables in the destination state. 
%%% SHORTEN BY LUDO
%If 
%$s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to $ then 
%% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
%		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
%		$\vars(g)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
%		$\vars(\codom(u))\!\subseteq\! \vars(s)\cup\vars(\alpha)\land 
%		\vars(\dom(u))\!\in\!\vars(s')$. %,  and $s= s'\Rightarrow J=\emptyset$. 
%\noteLH{is dom and codom clear?}
%\noteSB{Not sure: I understand, because I know what is meant, but this is not a canonical use, so should be explained. Will be solved if we roll back to \mbox{$(?x_i := e_i)^{i \in I}$} or \mbox{$\overline{?x := e}$}.}
%\noteLH{COORDINATION: I believe we should remove the constraints}		
%\end{itemize}
\end{definition}
%Now we define
%pNets as constructors for hierarchical behavioural structures.
A pNet composes several pNets, pLTSs, and holes.
A  pNet  exposes
 global actions resulting from the synchronisation of internal actions in some sub-pNets, and some actions of the holes. \newCoord{As holes are process parameters, synchronisation with a hole has an obvious concrete meaning when a process is put inside the hole and emits the action. We  also define a semantics for open pNets with holes where open transitions express the fact that a pNet can performs a transition provided one or several holes emit some actions.}
\newCoord{This synchronisation  is specified by  \emph{synchronisation vectors} expressing the synchronous interaction between actions inside sub-pNets and holes, data transmission is expressed classically using action parameters.}
Actions involved in the synchronisation vectors do
not need to distinguish input variables, \ie they 
have the form $a(\Expr_j^{j \in J})$.

@@

%\paragraph{Operational Semantics.}
%\label{section:op-semantics}
\renewcommand{\Pred}{g}
\renewcommand{\Post}{u}
The semantics of an open pNet is expressed as  an automaton where each transition coordinates the actions of several holes, 
the transition occurs if some predicates hold, and can involve  
 state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transition]
  \label{defn:OpenTransitions}
  An \emph{open transition} over a
  set of holes $J$  and a set of states $\mathcal{S}$ is 
       a structure of the form:
\\[-2ex]       
       \begin{mathpar}
       \inferrule*[myfraction=\reddottedrule]
       {       \beta_j^{j\in J}, \Pred, \Post}
       {s \OTarrow {{\alpha}}s'}
       \end{mathpar}
       Where $s, s'\in\mathcal{S}$ and $\beta_j\in\actions{\variables}$
        is an action of the hole $j$. $\alpha$ is the resulting global action. \Pred\ is a predicate 
       over the different variables of the
       terms, labels, and states  $\beta_j$, $s$, $\alpha$. $\Post\in 
       \assigns{\variables}$\ is 
       a set of assignments that are the effects of the transition.
Open transitions are identified modulo logical equivalence on their predicate.
%%   \noindent
%%   \begin{tabular}[b]{@{}p{0.86\columnwidth}c{}@{}}
%%     of holes $J$ and a set of states $\mathcal{S}$ is a structure
%%     of the form to the right,
%% %
%%     &
%% %
%%     %% \\[-2ex]	
%%     %% \begin{mathpar}
%%     \raisebox{1mm}[0pt]{
%%       $\inferrule*[myfraction=\reddottedrule]
%%       {	\beta_j^{j\in J}, \Pred, \Post}
%%       {s \OTarrow {{\alpha}}s'}$
%%       }
%%     %% \end{mathpar}
%% %
%%     \\
%% %
%%     \multicolumn{2}{p{\columnwidth}}{%
%% %
%%     where $s, s'\in\mathcal{S}$;
%%     $\beta_j\in\actions{\variables}$ is an action of the hole
%%     $j$; $\alpha$ is the resulting global action; {\Pred} is a
%%     predicate over the different variables of the terms, labels,
%%     and states $\beta_j$, $s$, $\alpha$; and $\Post\in
%%     \assigns{\variables}$ is a set of assignments that are the
%%     effects of the transition.  Open transitions are identified
%%     modulo logical equivalence on their predicate.
%% %
%%     }
%%   \end{tabular}
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

\newCoord{The red dotted rule somehow expresses an implication, stating that if actions of the holes perform the designed actions and the condition is verified, then the variables are modified and the state changes. This implication however  uses a simple logic with the 
 expressive power  given by the predicate algebra (it must include logical 
 propositions and equality). 
 Proposition and inference rules of the paper use a standard logic, while predicates inside the open transitions should use a more restricted logic, typically a logic that could  be handled mechanically and expressed by terms that can be encoded in a simple syntax.
Open transitions express in a symbolic way, transitions  that are not only parametrised with variables but also actions of not yet known processes.}

@@

\end{small}
It requires the hole at label $B$ to fire a $fail(b)$ action, with the condition $b=true$ and the state variable $t<Max$. In this case, the global pNet goes from state $\ostate{11}$ that has an internal variable $t$ to state $\ostate{14}$ where the state variable $t'$ has the value $t+1$. This emits a global action $\alpha(t)$. \newCoord{Figure~\ref{fig:ArchFailure:OA} illustrates a whole open automaton.}

@@

This section presents the extension of the theory of
  architectures \cite{AttieBBJS16-architectures-faoc} with data and
  briefly discusses a special case of priority models, called
  \emph{maximal progress}.

@@

In this paper, we  use  a refined
version of the Failure Monitor architecture
from~\cite{CubETH-case-study} as a running example.  Although \fig{schema:ArchFailure:BIP}
shows the full definition of this architecture, we will explain its
various elements progressively.
\fig{schema:ArchFailure:BIP} shows  components
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$(imer) and $\NameCtrl$(ontrol), with interfaces
$\btuple{
  \{t, z\},
  \{\PortTick, \PortCancel, \PortTOT\},
  \bigl\{\!\PortTick \mapsto \{t, z\}\!\bigr\}
}$,
and
$\btuple{
  \{\mathit{zone}\},
  \{\PortReset, \PortFail, \PortResume, \PortTOC\},
  \bigl\{\!\PortFail \mapsto \{\mathit{zone}\}\!\bigr\}
}$
%\noteEM{Attention, bug dans la figure 2, ou tu mets $reset\mapsto \mathit{zone}$}\noteSB{Done!}
respectively.
%
%% \[
%% \begin{array}{@{}c@{\hspace{3mm}}c@{}}
%%   \bigl(
%%   \{t, z\},
%%   \{\PortTick, \PortCancel, \PortTOT\},
%%   \export[\NameTimer]
%%   \bigr)\,,
%%   %% \qquad
%%   &
%%   \bigl(
%%   \{\mathit{zone}\},
%%   \{\PortReset, \PortFail, \PortResume, \PortTOC\},
%%   \export[\NameCtrl]
%%   \bigr)\,,
%%   \\
%%   \export[\NameTimer](p) =
%%   \begin{cases}
%%     \{t, z\}, & \text{if } p = \PortTick,\\
%%     \emptyset, & \text{otherwise,}
%%   \end{cases}
%%   %% \]
%%   %% and
%%   %% \[
%%   %% \quad
%%   &
%%   \export[\NameCtrl](p) =
%%   \begin{cases}
%%     \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
%%     \emptyset, & \text{otherwise.}
%%   \end{cases}
%% \end{array}
%% \]
%
%% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be an integer.  Variables $z$
and $\mathit{zone}$ are  integer intervals of
the form $[l,u]$.  Both types are implicitly equipped with a
natural ordering: increasing integers and interval inclusion.

@@

The use of the intermediate valuation $\tilde{\val{}{}}$ in the
conclusion and the third premise of  \eq{comp:semantics}
allows some  variables to get new values before the
transition is  fired.  Thus the component is \emph{open}
to the exchange of data with its environment.  However, the
fourth premise  states that only the variables exported through the ports participating to the interaction can be affected by the data transfer.
%restricts the set of
%variables that can be affected by such data transfer to those
%that are exported through the ports participating in the
%interaction.

@@

\begin{definition}[Architecture]
  \label{defn:arch}
  An \emph{architecture} is a tuple $A = \tuple{\cC, V_A, P_A,
    \export[A], \Gamma}$, where\vspace{-1.5ex}
  \begin{itemize}
\item
  $\cC$ is a finite set of \emph{coordinating components},
  %% with pairwise disjoint sets of ports and variables,
  such that
  $\bigcup_{C \in \cC} P_C \subseteq P_A$ and
  $\bigcup_{C \in \cC} V_C \subseteq V_A$;
%
  ports in $P_A \setminus \bigcup_{C \in \cC} P_C$, which do not
  belong to any of the coordinating components are called
  \emph{dangling};

@@

>>>>

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\Eg}[1][\ ]{E.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}
\newcommand{\Wrt}[1][\ ]{W.r.t.#1}
\newcommand{\resp}[1][\ ]{resp.#1}

@@

\newcommand{\order}{\leqslant}
\newcommand{\ordbool}{\ensuremath{\sB^{\order}}}
\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\signature}{\ensuremath{\Sigma}}
\newcommand{\variables}{\ensuremath{\cV}}
\newcommand{\Talg}{\ensuremath{\cT_{\signature,\variables}}}
\newcommand{\actions}[1]{\ensuremath{\mathit{Act}_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\monexprs}[1]{\ensuremath{\sE^{\order}_{#1}}}
\newcommand{\boolexprs}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\guards}[1]{\ensuremath{\ordbool_{#1}}}
\newcommand{\assigns}[1]{\ensuremath{\sA_{#1}}}
\newcommand{\updates}[1]{\ensuremath{\sA^{\order}_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\data^{#1}}}
\newcommand{\val}[3][]{\ensuremath{#1{\sigma}^{#2}_{#3}}}

%% CTL
\newcommand{\AG}[1][\ ]{\ensuremath{\mathtt{AG}#1}}
\newcommand{\AF}[1][\ ]{\ensuremath{\mathtt{AF}#1}}
\newcommand{\EF}[1][\ ]{\ensuremath{\mathtt{EF}#1}}
\newcommand{\EX}[1][\ ]{\ensuremath{\mathtt{EX}#1}}
\newcommand{\AW}[3][\ ]{\ensuremath{\mathtt{A}\left[#2\ \mathtt{W}\ #3\right]#1}}
\newcommand{\EU}[3][\ ]{\ensuremath{\mathtt{E}\left[#2\ \mathtt{U}\ #3\right]#1}}
\newcommand{\EW}[3][\ ]{\ensuremath{\mathtt{E}\left[#2\ \mathtt{W}\ #3\right]#1}}

@@

\begin{abstract}
We provide a solution for the design of safe concurrent systems by
compositional application of verified design patterns\mdash 
called \emph{architectures}\mdash to a small set of functional
components.  To this end, we extend the theory of architectures
developed previously for the BIP framework with the elements necessary
for handling data: definition and operations on data domains, syntax
and semantics of composition operators involving data transfer.
%
\newCoord{We provide a property preservation result, which states that
%
under
monotonicity assumptions on guards and assignments and, in presence of the \emph{maximal progress} special case of
  priority models, additional upwards compatibility assumptions on the architectures,} composition of
architectures preserves their characteristic safety properties.
To verify that individual architectures do enforce their associated
properties, we provide an encoding into open pNets, an intermediate
model  that supports SMT-based verification.
%% Additionally, the translation into  open pNet  \noteEM{ICI}{gives
%% us a strong insight and a comparison of the coordination mechanisms
%% provided by the two frameworks}.
The approach is illustrated by a case
study based on a previously developed BIP model of a nanosatellite
on-board software.

@@

We assume the existence of a universal data domain
given as a partially-ordered set $(\data, \order)$, potentially
encompassing several copies of any given data type with
different orders.  We assume that $(\data,
\order)$ comprises both the unordered set of Booleans $\sB =
\tuple{\{\true, \false\}, \emptyset}$ and the naturally ordered one
$\sB^\order = \tuple{\{\true, \false\}, \{\false \order \true\}}$, and
similarly for integer and real numbers; as well as the set of
intervals ordered by inclusion.
%In practice, and in the examples of
%this paper, all the variables are sorted, meaning that their values
%can only range over given subsets of $\data$, such as $\sB$ or
%$\sB^\order$.
%
When speaking of an ordered sort, \eg $\sB^\order$, we will assume that it
forms a meet-semilattice and denote by $\wedge$ the meet
operator.

@@

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
%\begin{itemize}
%\item
$S$ is a set of states; %$\vars(s)$ denotes the set of associated variables;
%\item
$s_0 \in S$ is the initial state;
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
%\item 
${\to} \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
% SB: I removed a new paragraph here, ok?
$\langle \alpha,~g,~u\rangle$,
where $\alpha$ is a parameterised action, $\alpha \in\actions{\variables}$; 
$g\in\boolexprs{\variables}$ is a guard over variables of the source state and the 
action, and $u\in\assigns{\variables}$
assigns updated value for variables in the destination state. 
%%% SHORTEN BY LUDO
%If 
%$s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to $ then 
%% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
%		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
%		$\vars(g)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
%		$\vars(\codom(u))\!\subseteq\! \vars(s)\cup\vars(\alpha)\land 
%		\vars(\dom(u))\!\in\!\vars(s')$. %,  and $s= s'\Rightarrow J=\emptyset$. 
%\noteLH{is dom and codom clear?}
%\noteSB{Not sure: I understand, because I know what is meant, but this is not a canonical use, so should be explained. Will be solved if we roll back to \mbox{$(?x_i := e_i)^{i \in I}$} or \mbox{$\overline{?x := e}$}.}
%\noteLH{COORDINATION: I believe we should remove the constraints}		
%\end{itemize}
\end{definition}
%Now we define
%pNets as constructors for hierarchical behavioural structures.
A pNet composes several pNets, pLTSs, and holes.
A  pNet  exposes
 global actions resulting from the synchronisation of internal actions in some sub-pNets, and some actions of the holes. \newCoord{As holes are process parameters, synchronisation with a hole has an obvious concrete meaning when a process is put inside the hole and emits the action. We  also define a semantics for open pNets with holes where open transitions express the fact that a pNet can performs a transition provided one or several holes emit some actions.}
\newCoord{This synchronisation  is specified by  \emph{synchronisation vectors} expressing the synchronous interaction between actions inside sub-pNets and holes, data transmission is expressed classically using action parameters.}
Actions involved in the synchronisation vectors do
not need to distinguish input variables, \ie they 
have the form $a(\Expr_j^{j \in J})$.

@@

%\paragraph{Operational Semantics.}
%\label{section:op-semantics}
\renewcommand{\Pred}{g}
\renewcommand{\Post}{u}
The semantics of an open pNet is expressed as  an automaton where each transition coordinates the actions of several holes, 
the transition occurs if some predicates hold, and can involve  
 state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transition]
  \label{defn:OpenTransitions}
  An \emph{open transition} over a
  set of holes $J$  and a set of states $\mathcal{S}$ is 
       a structure of the form:
\\[-2ex]       
       \begin{mathpar}
       \inferrule*[myfraction=\reddottedrule]
       {       \beta_j^{j\in J}, \Pred, \Post}
       {s \OTarrow {{\alpha}}s'}
       \end{mathpar}
       Where $s, s'\in\mathcal{S}$ and $\beta_j\in\actions{\variables}$
        is an action of the hole $j$; $\alpha$ is the resulting global action; {\Pred} is a predicate 
       over the different variables of the
       terms, labels, and states  $\beta_j$, $s$, $\alpha$. $\Post\in 
       \assigns{\variables}$\ is 
       a set of assignments that are the effects of the transition.
Open transitions are identified modulo logical equivalence on their predicate.
%%   \noindent
%%   \begin{tabular}[b]{@{}p{0.86\columnwidth}c{}@{}}
%%     of holes $J$ and a set of states $\mathcal{S}$ is a structure
%%     of the form to the right,
%% %
%%     &
%% %
%%     %% \\[-2ex]	
%%     %% \begin{mathpar}
%%     \raisebox{1mm}[0pt]{
%%       $\inferrule*[myfraction=\reddottedrule]
%%       {	\beta_j^{j\in J}, \Pred, \Post}
%%       {s \OTarrow {{\alpha}}s'}$
%%       }
%%     %% \end{mathpar}
%% %
%%     \\
%% %
%%     \multicolumn{2}{p{\columnwidth}}{%
%% %
%%     where $s, s'\in\mathcal{S}$;
%%     $\beta_j\in\actions{\variables}$ is an action of the hole
%%     $j$; $\alpha$ is the resulting global action; {\Pred} is a
%%     predicate over the different variables of the terms, labels,
%%     and states $\beta_j$, $s$, $\alpha$; and $\Post\in
%%     \assigns{\variables}$ is a set of assignments that are the
%%     effects of the transition.  Open transitions are identified
%%     modulo logical equivalence on their predicate.
%% %
%%     }
%%   \end{tabular}
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

\newCoord{The red dotted rule expresses the implication stating that if the holes perform the designated actions and the condition $g$ is verified, then the variables are modified and the state changes. This implication however  uses a simple logic with the 
 expressive power  given by the predicate algebra (it must include logical 
 propositions and equality). 
 Proposition and inference rules of the paper use a standard logic, while predicates inside the open transitions should use a more restricted logic, typically a logic that could  be handled mechanically and expressed by terms that can be encoded in a simple syntax.
Open transitions express in a symbolic way, transitions  that are not only parametrised with variables but also actions of not yet known processes.}

@@

\end{small}%
It requires the hole at label $B$ to fire a $fail(b)$ action, with the condition $b=true$ and the state variable $t<Max$. In this case, the global pNet goes from state $\ostate{11}$ that has an internal variable $t$ to state $\ostate{14}$ where the state variable $t'$ has the value $t+1$. This emits a global action $\alpha(t)$. \newCoord{Figure~\ref{fig:ArchFailure:OA} illustrates a complete open automaton.}

@@

This section presents the extension of the theory of
  architectures \cite{AttieBBJS16-architectures-faoc} with data and
  briefly discusses a special case of priority models, called
  \emph{maximal progress}.
%
  \newCoord{These extensions require us to define the framework in a
    manner that would allow formulating and proving the property
    preservation result (\thm{combining} below).  In
    \cite{AttieBBJS16-architectures-faoc}, this result is obtained by
    requiring, in the definition of architecture composition, that an
    interaction among coordinated components be only possible if both
    architectures ``agree'' that it should be enabled.  With respect
    to data, the main difficulty lies in ensuring that this
    ``agreement'' extends to the transferred data values.  A trivial
    extension would allow an interaction only if the data values
    proposed by both architectures coincide.  Considering this
    requirement too restrictive, we go beyond by assuming the data
    domains to be ordered and taking the \emph{meet} of the proposed
    values.  The property preservation result then holds independently
    of the proposed values, provided that guards and update
    assignments are monotonic.

    An important insight is that, although the requirement that guards
    and update assignments be monotonic appears to be a limitation, it
    is, in fact, a generalisation of the usual setting.  Indeed, the
    usual settings, where data domains are not ordered, can be
    recovered here by considering trivial partial orders with no two
    distinct elements being comparable.  In such case, \emph{all
      expressions} are trivially monotonic.

    The intuition behind the proof of the preservation of safety
    properties in \cite{AttieBBJS16-architectures-faoc} is simple.
    The composition of two architectures combines the ``constraints''
    that they impose on the possible executions of the system: as
    stated above, an interaction is only enabled if both architectures
    ``agree''.  In \cite{Baranov17-phd}, it is shown that this
    intuition extends well to priorities in the \emph{offer semantics}
    of BIP.  However, this is not the case in the classical semantics.
    In this section, we informally discuss the special case of the
    maximal progress priority models, where property preservation does
    hold in the classical semantics of BIP.
%
}

@@

In this paper, we  use  a refined
version of the Failure Monitor architecture
from~\cite{CubETH-case-study} as a running example.  Although \fig{schema:ArchFailure:BIP}
shows the full definition of this architecture, we will explain its
various elements progressively.
\fig{schema:ArchFailure:BIP} shows  components
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$(imer) and $\NameCtrl$(ontrol), with interfaces
$\btuple{
  \{t, z\},
  \{\PortTick, \PortCancel, \PortTOT\},
  \bigl\{\!\PortTick \mapsto \{t, z\}\!\bigr\}
}$,
and
$\btuple{
  \{\mathit{zone}\},
  \{\PortReset, \PortFail, \PortResume, \PortTOC\},
  \bigl\{\!\PortFail \mapsto \{\mathit{zone}\}\!\bigr\}
}$
%\noteEM{Attention, bug dans la figure 2, ou tu mets $reset\mapsto \mathit{zone}$}\noteSB{Done!}
respectively.
%
%% \[
%% \begin{array}{@{}c@{\hspace{3mm}}c@{}}
%%   \bigl(
%%   \{t, z\},
%%   \{\PortTick, \PortCancel, \PortTOT\},
%%   \export[\NameTimer]
%%   \bigr)\,,
%%   %% \qquad
%%   &
%%   \bigl(
%%   \{\mathit{zone}\},
%%   \{\PortReset, \PortFail, \PortResume, \PortTOC\},
%%   \export[\NameCtrl]
%%   \bigr)\,,
%%   \\
%%   \export[\NameTimer](p) =
%%   \begin{cases}
%%     \{t, z\}, & \text{if } p = \PortTick,\\
%%     \emptyset, & \text{otherwise,}
%%   \end{cases}
%%   %% \]
%%   %% and
%%   %% \[
%%   %% \quad
%%   &
%%   \export[\NameCtrl](p) =
%%   \begin{cases}
%%     \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
%%     \emptyset, & \text{otherwise.}
%%   \end{cases}
%% \end{array}
%% \]
%
%% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be \newCoord{of type Integer (with trivial ordering).  Variables $z \bydef{=} [z.l, z.u]$
and $\mathit{zone} \bydef{=} [\mathit{zone}.l, \mathit{zone}.u]$ are   of type Integer Interval 
ordered by interval inclusion}.

@@

The use of the intermediate valuation $\tilde{\val{}{}}$ in the
conclusion and the third premise of  \eq{comp:semantics}
allows some  variables to get new values before the
transition is  fired.  Thus the component is \emph{open}
to the exchange of data with its environment.  However, the
fourth premise  states that only the variables exported through the ports participating in the interaction can be affected by the data transfer.
%restricts the set of
%variables that can be affected by such data transfer to those
%that are exported through the ports participating in the
%interaction.

@@

\begin{definition}[Architecture]
  \label{defn:arch}
  An \emph{architecture} is a tuple $A = \tuple{\cC, V_A, P_A,
    \export[A], \Gamma}$, where\vspace{-1.5ex}
  \begin{itemize}
  \item $P_A$ and $V_A$ are sets of ports and variables, respectively;
\item
  $\cC$ is a finite set of components, which we call \emph{coordinators},
  %% with pairwise disjoint sets of ports and variables,
  such that
  $\bigcup_{C \in \cC} P_C \subseteq P_A$ and
  $\bigcup_{C \in \cC} V_C \subseteq V_A$;
%
  ports in $P_A \setminus \bigcup_{C \in \cC} P_C$, which do not
  belong to any of the coordinators are called
  \emph{dangling};

@@

\newCoord{Below, we present both the encodings of coordinators and
  interaction models.  The key constraint is that we encode each
  connector by one synchronisation vector.  This is necessary to
  1)~preserve the structure of the system and 2)~allow the encoding of
  maximal progress.

  Although somewhat technical, the encoding of coordinators is, in
  fact, pretty straightforward, comprising three key ideas: 1)~we
  introduce an additional transition (hence also an additional state)
  to explicitly initialise the variables; 2)~we introduce additional
  input variables to manipulate the values provided to the coordinator
  by the rest of the system for all exported variables; and
  3)~following the classical technique~\cite{milner83-calculi}, we
  simulating absence of action by an additional loop transition.
